# Aprendizaje semi-supervisado

En esta sección estudiaremos un tipo de problema intermedio entre supervisado y no supervisado, que es muy común en la práctica.  

## Librería `s2net`

La librería `s2net` de R resuelve un problema de optimización similar al elastic-net, pero para tratar con datos semi-supervisados. No obstante, es más general, e incluye el caso de datos supervisados también. Fue concebido específicamente para resolver problemas de transfer-learning en datos de alta dimensión, donde además de ajustar un modelo lineal, se desea incluir solamente un subconjunto de variables explicativas en el modelo (como hace elastic-net).

Más información aquí: https://github.com/jlaria/s2net

You can install the released version of `s2net` from [CRAN](https://CRAN.R-project.org) with:

``` r
install.packages("s2net")
```

The development version can be installed with:

``` r
devtools::install_github("jlaria/s2net", build_vignettes = TRUE)
```

This is a basic example which shows you how to use the package. Detailed examples can be found in the documentation and vignettes. 

```{r}
library(s2net)
# Auto-MPG dataset is included for benchmark
data("auto_mpg")
```

Semi-supervised data is made of a labeled dataset `xL`, the labels `yL`, and unlabeled data `xU`.  Package `s2net` includes the function `s2Data` to process semi-supervised datasets. 

```{r}
head(auto_mpg$P2$xL, 2) # labeled data
head(auto_mpg$P2$yL, 2) # labels
head(auto_mpg$P2$xU, 2) # unlabeled data

train = s2Data(auto_mpg$P2$xL, auto_mpg$P2$yL, auto_mpg$P2$xU, preprocess = TRUE)

head(train$xL, 2)
```

The data is centered and scaled, and factor variables are automatically converted to numerical dummies. Constant columns are also removed. If we wanted to use validation/test data, we must pre-process it according to the training data, with:

```{r}
valid = s2Data(auto_mpg$P2$xU, auto_mpg$P2$yU, preprocess = train)
```

There are two ways to fit a semi-supervised elastic-net using `s2net`. The easiest way is using the function `s2netR`, that returns an object of S3 class `s2netR`.

```{r}
model = s2netR(train, 
               params = s2Params(lambda1 = 0.01, 
                                 lambda2 = 0.01, 
                                 gamma1 = 0.01, 
                                 gamma2 = 100, 
                                 gamma3 = 0.1))

class(model)
model$beta

ypred = predict(model, valid$xL)
```
